\subsection*{Data Sets}
\textbf{cifar-10-batches-py} - CIFAR-10 image data batches. We only use the test batch\\
\textbf{cifar10h-raw.csv, cifar10h.csv} - Raw and processed CIFAR-10H dataset. \\
\textbf{Y\_cifar10h.csv} - Aggregated vote distribution for the 10'000 images of the CIFAR-10H dataset. \\

\textbf{theta.csv, theta\_*.csv} - Estimated $\hTheta$ for the different models. \\
\textbf{pi.csv, pi\_*.csv} - Estimated $\hbpi$ values for the different models. \\
\textbf{tau.csv, tau\_*.csv} - Estimated $\hbtau$ values for the different models. \\
\textbf{theta\_old.csv, theta\_old\_*.csv} - Estimated $\hbvtheta$ for the different models. \\
\textbf{variance\_theta.csv, variance\_theta\_*.csv} - Estimated $\widehat{\V}(\hbvtheta)$ values for the different models. \\
\textbf{post\_entropy.csv, post\_entropy\_*.csv} - Estimated posterior entropy $H_{post}$ values for the different models. \\

The \textbf{*} in the file names represents the different models. The models are: \\
\textbf{L} - Slow reaction time. \\
\textbf{R} - Fast reaction time. \\
\textbf{S} - Subcategories based on reaction time. \\
\textbf{Se} - Subcategories based on entropy. \\
No \textbf{*} - Model on full data set.

\textbf{top10\_entropy.csv} - File names of the images with the top 10 entropy. \\
\textbf{special\_entropy.csv} - File names of the images with the similar vote distribution but different entropy. \\


\subsection*{R Files}
\textbf{Data\_Cleaning.R} - Script for data preparation.\\
\textbf{Descriptive\_Statistics.R} - Code for creating the graphics for the chapter data. \\
\textbf{Image\_Ambiguity.R} - Code for the results in the chapter on image ambiguity. \\
\textbf{Subcategory\_Ambiguity.R} - Code for the results in the chapter on subcategory ambiguity. \\
\textbf{Reaction\_Time\_Ambiguity.R} - Code for the results in the chapter on reaction time effect. \\
\textbf{Model\_RT.R} - Code for the reaction time GLMM. \\

\subsection*{Python Files}
\textbf{model\_funcs.py} - Script for data preparation and sem algorithm\\
\textbf{cifar\_images.ipynb} - Script for extracting the images from the CIFAR-10 dataset.\\
\textbf{Image\_Ambiguity.ipynb} - SEM Code for the results in the chapter on image ambiguity. \\
\textbf{Subcategory\_Ambiguity.ipynb} - SEM Code for the results in the chapter on subcategory ambiguity. \\
\textbf{Reaction\_Time\_Ambiguity.ipynb} - SEM Code for the results in the chapter on reaction time effect. \\