We assume that we have a setting, in which we have images that are to be
labelled. The annotator has to choose the label from a rather small set of
predefined labels. We assume that the annotator is not perfect and that there is
some uncertainty in the labelling process. This uncertainty can arise from
various sources such as ambiguity in the data, subjective interpretation by the
annotator, and noise in the labelling process. We further assume that there is
an unknown underlying ground truth to every image. These scenarios might arise in
real-world applications, such as
\begin{itemize}
    \item medical imaging: classifying images of X-rays or MRIs into different medical conditions
    \item wildlife monitoring: classifying images of animals into different species
    \item quality control in manufacturing: classifying images of products into different quality categories
    \item agriculture: classifying images of crops into different growth or health stages
    \item environmental monitoring: classifying images of landscapes into different local climate zones
\end{itemize}

We base our methodology on the work of
\cite{hechingerCategorisingWorldLocal2024}, who used a Bayesian Latent Class
Mixture Model to quantify the labelling uncertainty. The estimation of
the parameters is done using a Stochastic Expectation-Maximization
algorithm.

\subsection{Bayesian Latent Class Mixture Model}
\label{BLCMM}

% Latent classes are subgroups within a population that are not directly observed. In our
% case, the population is the set of all images and the latent classes are the
% true categories of the images. Each image belongs to exactly one latent class, which
% is (assumed to be) unknown. Furthermore, we assume that the observed given labels come
% from a mixture of the different distributions, based on the latent class of the
% image. This means that each class has its own parameters that determine the
% probability of observing a certain label, given the latent class. Using bayesian
% methods, we can estimate the parameters of the models by specifying prior
% distributions for the parameters and use the observed data to update the
% prior distributions to posterior distributions.\bigskip

Let $i$ with $i = 1, \ldots, n$ be the index of the images and $j$ with $j = 1,
\ldots, J$ be the index of the annotators. We denote the label that the
annotator $j$ assigns to the image $i$ by $V^{(i)}_j$, with $V^{(i)}_j \in \{1,
\ldots, K\}$, where $K$ is the number of classes. This leads to the definition
of the accumulated votes of all annotators per image $i$ as $\bY^{(i)} :=
(Y_1^{(i)}, \ldots, Y_K^{(i)})$, with $Y_k^{(i)} := \sum_{j=1}^{J^{(i)}}
I(V^{(i)}_j = k)$, $I(\cdot)$ being the indicator function and $J^{(i)}$ being
the number of annotators that labelled the image $i$. We will refer to
$\bY^{(i)}$ as the vote distribution of image $i$.\bigskip

We assume that each image $i$ comes from a latent class, i.e. the image $i$
belongs to a true category $Z^{(i)}$, with $Z^{(i)} \in \{1, \ldots, K\}$ which
is unknown (latent) and define $\bZ^{(i)}$ as the one-hot encoded vector of
$Z^{(i)}$, i.e. $\bZ^{(i)} = (Z_1^{(i)}, \dots, Z_K^{(i)}) = (I(Z^{(i)} = 1), \ldots, I(Z^{(i)} = K))$. Using
this notation, we formulate our model as follows:
\begin{align}
    \bZ^{(i)} &\sim \text{Multi}(\bpi, 1), i.i.d. \quad \text{with} \quad \bpi = (\pi_1, \ldots, \pi_K) \quad \text{and} \quad \sum_{k=1}^K \pi_k = 1, \label{eq:1} \\
    \bY^{(i)} | Z^{(i)} &\sim \text{Multi}(\btheta_{Z^{(i)}}, J^{(i)}) \quad \text{with} \quad \btheta_z = (\theta_{z1}, \ldots, \theta_{zK}) \quad \text{and} \quad \sum_{k=1}^K \theta_{zk} = 1. \label{eq:2}
\end{align}
We call $\pi_k$ the prior probability of class $k$, i.e. the probability that an
image belongs to class $k$ without knowing anything about the image. The vector
$\btheta_z$ contains the probabilities $\theta_{zk}, k = 1,\dots,K$ of the
annotators assigning the label $k$ to an image of class $z$. We can write the
$\btheta_k$ vectors as rows of a matrix $\Theta$, where $\Theta := (\btheta_1,
\ldots, \btheta_K) = 
\begin{pmatrix}
    \horzbar & \btheta_1 & \horzbar \\
             & \vdots    &          \\
    \horzbar & \btheta_K & \horzbar \end{pmatrix}$, which we call the confusion
matrix. In Figure \ref{fig:confusion_matrix} we have seen an example of this
confusion matrix. Given a latent class $Z^{(i)}$ and the confusion matrix
$\Theta$, we have a specific distribution for the votes $\bY^{(i)}$, i.e. the
observed labels come from a mixture of different distributions (mixture model),
based on the latent class of the image $i$. We obtain the likelihood contribution of
the image $i$ through
\begin{equation}
    \P(\bY^{(i)}; \bpi, \Theta) = \sum_{z=1}^K \P(\bY^{(i)} | Z^{(i)} = z; \Theta) \P(Z^{(i)} = z; \bpi) = \sum_{z=1}^K \pi_z \P(\bY^{(i)}; \btheta_z).
\end{equation}
In order to estimate the parameters $\pi$ and $\Theta$, we use the Stochastic
Expectation Maximization algorithm, which is described in the following section.

\subsection{Stochastic Expectation Maximization Algorithm}
\label{SEM}

Using the Expectation Maximization algorithm (\cite
{dempsterMaximumLikelihoodIncomplete1977}), we can estimate the parameters of
the model. The EM algorithm is an iterative method that alternates between the
E-step and the M-step. In the E-step, we compute the expected value of
$\bZ^{(i)}$ given the observed data $\bY^{(i)}$ and the current estimate of the
parameters $\hbpi$ and $\hTheta$. This results in a complete dataset that
can then be used to estimate the best parameters in the M-step. The new estimates
of the parameters are then used in the next iteration of the E-step.

The stochastic expectation maximization (SEM) algorithm
(\cite{celeuxStochasticVersionsEm1996}) is a stochastic version of the EM
algorithm, which is especially useful when one is also interested in the
uncertainty of the estimates. The SEM algorithm is based on the idea of the EM
algorithm, but instead of computing the expected value of $\bZ^{(i)}$, it
samples $\bZ^{(i)}$ from the posterior distribution.\bigskip

The SEM algorithm works as follows:
\begin{enumerate}
    \item Initialize the parameters $\hbpi_0$ and $\hTheta_0$. If no prior
    knowledge is available, one can initialize $\hbpi_0 = (\frac{1}{K}, \ldots, \frac{1}{K})$ and
    draw $\hTheta_0$ randomly from a Dirichlet distribution with parameters $K$ and $\alpha =
    (2K, \ldots, 2K)$.

    \item Repeat for $t = 0, \ldots, T$:
    \begin{enumerate}
        \item E-Step: Compute
        \begin{align}
            \htau_{z,t}^{(i)} = \P(Z_{t}^{(i)} = z|
        \bY^{(i)}; \hbpi_{t}, \hTheta_{t}) = \frac{\hbpi_{z,t} \P(\bY^{(i)};
        \hbtheta_{z,t})}{\sum_{z' = 1}^K \hbpi_{z',t} \P(\bY^{(i)};
        \hbtheta_{z',t})}. \label{eq:tau}
        \end{align}
        Use the $\hbtau_t^{(i)}$ to randomly sample
        $\bZ_{t}^{(i)} ~ \sim \text{Multi}(\hbtau_t^{(i)}, 1)$ for all $i = 1,\dots, n$.
        
        \item Update the parameter estimates $\hbpi$ and $\hTheta$ using the
        sampled $\bZ^{(i)}$. We use the maximum likelihood estimator (MLE) for
        $\hbpi$ and $\hTheta$ to update the estimates. The MLE for $\hbpi$ is
        given by
        \begin{align}
            \hbpi_{t+1} &= \frac{1}{n} \sumin \bZ_k^{(i)} \label{eq:pi_mle} \\
            \htheta_{zk,t+1} &= \frac{\sumin I(Z_t^{(i)} = z) Y_k^{(i)}}{\sumin I(Z_t^{(i)} = z) J^{(i)}}.\label{eq:theta_mle}
        \end{align}
    \end{enumerate}
    \item Compute the final estimates $\hbpi, \hTheta$ and $\hbtau$, by
        averaging over the samples. Note that we discard the first $t_0$ samples
        to avoid the burn-in phase, where the algorithm has not yet converged to
        the true distribution.
        \begin{align}
            \hbpi &= \frac{1}{T-t_0} \sum_{t=t_0 + 1}^T \hbpi_t \\
            \hTheta &= \frac{1}{T-t_0} \sum_{t=t_0 + 1}^T \hTheta_t \\
            \hbtau^{(i)} &= \frac{1}{T-t_0} \sum_{t=t_0 + 1}^T \hbtau_t^{(i)}.
        \end{align}
\end{enumerate}

The equations above can be interpreted as follows: 
\begin{itemize}
    \item Equation \ref{eq:tau} computes the posterior probability that
    the image $i$ belongs to class $z$, given the observed votes $\bY^{(i)}$ and
    current estimates of the parameters $\hbpi_t$ and $\hTheta_t$.
    \item Equation \ref{eq:pi_mle} computes the relative frequency of the latent
    classes.
    \item Equation \ref{eq:theta_mle} computes how many images of latent class
    $z$ have been labelled with label $k$, compared to the total number of votes
    that the latent class $z$ received. This can also be expressed in matrix
    notation, see Appendix \ref{app} equation \ref{eq:theta_mle_matrix}.
\end{itemize}

Mixture models usually have an identifiability problem, i.e. the latent classes
are not uniquely identifiable. This means that we don't know which latent class
$z = 1,\dots,K$ corresponds to which label $k \in \{\text{airplane},
\text{automobile}, \dots, \text{truck}\}$. To solve this problem, we can use Bayes
Theorem to obtain $$\P(Z^{(i)} = z| V^{(i)} = k) \propto \P(V^{(i)} = k| Z^{(i)} = z)
\P(Z^{(i)} = z)$$ and use the estimated confusion matrix $\hTheta$ and the
estimated latent class probabilities $\hbpi$ to compute the posterior
probabilities $\P(Z^{(i)} = z| V^{(i)} = k)$ for all $z = 1,\dots,K$ and $k =
1,\dots,K$. Based on that, we can then assign the labels to the latent classes
and permute the rows of the confusion matrix $\hTheta$ for better interpretability.\bigskip

% In our case, it is sufficient to sort the rows of $\hTheta$ such that the highest
% value in each row is on the diagonal. In generell, the label switching problem is
% a common problem in Bayesian Mixture Models, where the labels of the latent
% classes are interchangeable.\bigskip


Using the chain of estimates $(\hbpi_t)_{t = t_0,\dots,T}$, $(\hTheta_t)_{t =
t_0,\dots,T}$ and $(\hbtau_t)_{t = t_0,\dots,T}$ and the samples of the latent
classes $(\bZ_t^{(i)})_{t = t_0,\dots,T}$ that are produced by the SEM
algorithm, we can estimate the uncertainty of the estimates. We are especially
interested in the uncertainty of the confusion matrix $\hTheta$. For that we
rewrite the matrix $\hTheta$ as a vector $\hbvtheta$ as defined below. Note
that the matrix $\hTheta$ is of full rank, because the rows of $\hTheta$ are
probability vectors that sum up to one. To account for this constraint, we
drop the last column of $\hTheta$ and denote the resulting matrix as $\hTheta_{-K}$
and its rows as $\hbtheta_{z, -K}$ for $z = 1,\dots,K$. Using this notation, we
can define the vector $\hbvtheta$ as
\begin{equation}
    \hbvtheta = (\hbtheta_{1, -K}, \ldots, \hbtheta_{K, -K})^\top = (\htheta_{1, 1}, \ldots, \htheta_{1, K-1}, \ldots, \htheta_{K, 1}, \ldots, \htheta_{K, K-1})^\top.
\end{equation}

The estimate of the covariance matrix of $\hbvtheta$, denoted by
$\widehat{\V}(\hbvtheta)$, is given below. The formulas are derived in
\cite{littleBayesMultipleImputation2002}, chapter 10.2. and
\cite{hechingerCategorisingWorldLocal2024}.

\begin{align}
    &\widehat{\V}(\hbtheta_{z, -K} | Z) = \frac{diag(\hbtheta_{z, -K}) - \hbtheta_{z, -K} \cdot \hbtheta_{z, -K}^\top}{\sumin I(Z^{(i)} = z)}\\
    &V_t := blockdiag(\widehat{\V}(\hbtheta_{1, -K} | Z_t), \ldots, \widehat{\V}(\hbtheta_{K, -K} | Z_t)) \\ \label{eq:V_t}
    &\widehat{\V}(\hbvtheta) = \frac{1}{T-t_0} \sum_{t=t_0 + 1}^T V_t + \frac{1}{T-t_0-1} \sum_{t=t_0 + 1}^T (\hbvtheta_t - \hbvtheta)(\hbvtheta_t - \hbvtheta)^\top
\end{align}

Note that given the true latent classes $Z$, the probability vectors
$\hbtheta_{z, -K}$ and $\hbtheta_{z', -K}$ are independent for $z \neq z'$. Hence
the matrix $V_t$ in equation \ref{eq:V_t} is block-diagonal.\bigskip

Using this methodology, we can quantify the uncertainty of the
votes and perform tests on the confusion matrix $\hTheta$. We can
also compute the posterior probabilities of the latent classes and use
those to compute a posterior entropy for each image, which quantifies the
ambiguity of the image.



