---
title: "cifar10h_DescriptiveAnalysis"
format: html
editor: visual
---

```{r setup, include=FALSE}
library(readr) # for reading csv files
library(knitr) # for rendering markdown
library(tidyverse) # for data manipulation and visualization
library(kableExtra) # for better table display
library(gt) # for better table display)

# Set echo, warning and message to FALSE
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

theme_set(theme_bw())

cifar10h <- read_csv("../cifar-10h/data/cifar10h-raw.csv")
```

```{r}
# transform the classes of the columns
cifar10h <- cifar10h %>%
  mutate(
    annotator_id = as.character(annotator_id),
    is_attn_check = factor(is_attn_check, levels = c(0, 1)),
    true_label = factor(true_label, levels = c(0:9)),
    chosen_label = factor(chosen_label, levels = c(0:9))
  ) %>%
  select(-c(cifar10_test_test_idx, image_filename))
```

# Data Description

```{r}
# Descriptive statistics
cifar10h %>%
  sample_n(10) %>%
  arrange(as.numeric(annotator_id)) %>%
  gt() %>% 
  tab_header(
    title = "CIFAR-10h Dataset",
    subtitle = "Random 10 rows of the CIFAR-10h dataset"
  ) %>%
  cols_label(
    annotator_id = md("**annotator_id**,<br>chr"),
    trial_index = md("**trial_index**,<br>int"),
    is_attn_check = md("**is_attn_check**,<br>boolean"),
    true_category = md("**true_category**,<br>fac\\w10lvl"),
    chosen_category = md("**chosen_category**,<br>fac\\w10lvl"),
    true_label = md("**true_label**,<br>fac\\w10lvl"),
    chosen_label = md("**chosen_label**,<br>fac\\w10lvl"),
    correct_guess = md("**correct_guess**,<br>boolean"),
    subcategory = md("**subcategory**,<br>chr"),
    reaction_time = md("**reaction_time**,<br>dbl"),
    time_elapsed = md("**time_elapsed**,<br>dbl")
  )
```

# Summary Statistics

## Distribution of the labels

```{r}
cifar10h %>%
  ggplot(aes(x = chosen_category)) +
  geom_bar() +
  geom_hline(yintercept = unique(table(cifar10h$true_category)), color = "blue") +
  scale_y_continuous(breaks = c(0, 20000, 40000, unique(table(cifar10h$true_category))),
                     labels = c(0, 20, 40, round(unique(table(cifar10h$true_category)), -3)/1000)) +
  labs(
    x = "Chosen Category",
    y = "Frequency (in thousands)"
  )

prop.table(table(cifar10h$true_category, cifar10h$chosen_category), margin = 1) %>%
  as.data.frame() %>%
  ggplot(aes(x = Var2, y = Var1, fill = sqrt(Freq))) +
  geom_tile() +
  scale_fill_gradient(low = "white", high = "orange") +
  labs(
    y = "True Category",
    x = "Chosen Category",
    fill = "Proportion"
  ) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  # add the value of the proportion on the tile
  geom_text(aes(label = scales::percent(round(Freq, 3))), size = 3) +
  theme(legend.position = "none")
```

On average the annotators correctly guessed the category of the image `r round(mean(cifar10h$correct_guess) * 100, 2)`\% of the time. The most common category chosen by the annotators was the true category of the image. The proportion of correct guesses for each category is shown in the heatmap above. The diagonal of the heatmap shows the proportion of correct guesses for each category.

## Reaction Time

```{r}
rt_med <- median(cifar10h$reaction_time / 1000)


cifar10h %>%
  ggplot(aes(x = reaction_time/1000)) +
  stat_ecdf() +
  scale_x_continuous(limits = c(0, 6), labels = scales::comma, breaks = c(seq(0, 6, 2), rt_med)) +
  geom_segment(aes(x = rt_med, y = 0, xend = rt_med, yend = 0.5), linetype = "dashed") +
  geom_segment(aes(x = 0, y = 0.5, xend = rt_med, yend = 0.5), linetype = "dashed") +
  labs(
    x = "Reaction Time (in seconds)",
    y = "Cumulative Proportion"
  )

cifar10h %>%
  ggplot(aes(x = reaction_time/1000)) +
  stat_density() +
  scale_x_continuous(limits = c(0, 6), labels = scales::comma, breaks = c(seq(0, 6, 2), rt_med)) +
  labs(
    x = "Reaction Time (in seconds)",
    y = "Cumulative Proportion"
  )



ggplot(data.frame(rt_t = transformed_data), aes(x = rt_t)) +
  stat_density() +
  #scale_x_continuous(limits = c(0, 6), labels = scales::comma, breaks = c(seq(0, 6, 2), rt_med)) +
  labs(
    x = "Reaction Time (in seconds)",
    y = "Cumulative Proportion"
  ) +
  stat_function(fun = dnorm, args = list(mean = mean(transformed_data), sd = sd(transformed_data)), color = "red") 

df <- data.frame(transformed_data = transformed_data, original_data = reaction_time)

# Create the ggplot
ggplot(df, aes(x = transformed_data, y = original_data)) +
  geom_point() +
  labs(x = "Transformed Reaction Time", y = "Original Reaction Time") +
  scale_x_continuous(labels = function(x) round(inverse_boxcox(x, lambda), 2))


```


```{r}
# Assuming your data is in a variable called reaction_time
reaction_time <- cifar10h %>%
  filter(reaction_time > 500, reaction_time < 6000) %>%
  mutate(reaction_time = reaction_time / 1000) %>%
  pull(reaction_time)


# Apply the Box-Cox transformation
boxcox_result <- boxcox(reaction_time ~ 1, plotit = TRUE)

# Extract the optimal lambda value
lambda <- boxcox_result$x[which.max(boxcox_result$y)]

# Apply the transformation with the optimal lambda
transformed_data <- (reaction_time^lambda - 1) / lambda

# Print the optimal lambda
print(paste("Optimal lambda:", lambda))

inverse_boxcox <- function(transformed_data, lambda) {
  if (lambda == 0) {
    return(exp(transformed_data))
  } else {
    return((lambda * transformed_data + 1)^(1 / lambda))
  }
}

# Histogram of the transformed data
hist(inverse_boxcox(transformed_data, lambda), breaks = 50, main = "Histogram of Box-Cox Transformed Data", xlab = "Transformed Reaction Time")

# Q-Q plot of the transformed data
qqnorm(transformed_data)
qqline(transformed_data, col = "blue")

```

